---
title: "Analyzing *Acropora* spp. mortality in the 2023 heatwave in Florida"
author: "Ross Cunning and Rich Karp"
output: html_document
date: "2024-01-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, echo = TRUE, cache = TRUE)
```

```{r load_libraries}
library(lme4)
library(emmeans)
library(tidyverse)
library(cowplot)
library(drc)
```

# Import processed dataset
```{r import_data}
# Processed data with sites for each data provider
df1 <- read_csv("data/processed/FLAcro23_data_processed_sites.csv") %>%
  mutate(subregion = case_when(subregion == "Tortugas--Dry Tortugas NP" ~ "Dry Tortugas", TRUE ~ subregion)) %>%
  mutate(subregion = factor(subregion, levels = c("Broward-Miami", "Biscayne", "Upper Keys",
                                                  "Middle Keys", "Lower Keys", "Dry Tortugas")))
# Processed data with sites clustered based on proximity
df <- read_csv("data/processed/FLAcro23_data_processed.csv") %>%
  mutate(subregion = case_when(subregion == "Tortugas--Dry Tortugas NP" ~ "Dry Tortugas", TRUE ~ subregion)) %>%
  mutate(subregion = factor(subregion, levels = c("Broward-Miami", "Biscayne", "Upper Keys",
                                                  "Middle Keys", "Lower Keys", "Dry Tortugas")))
```

### Map dataset in space and time
```{r map_data}
# How many individual sites
n_distinct(df1$site)

# Get the maximum number of ramets of each species ever surveyed (on one date) at a siteclust
# Get the sum of max total counted in each subregion
## This could underestimate, because diff sites in same siteclust surveyed on same date would be lost
dfs <- df %>%
  group_by(siteclust, lat, lon, subregion, species, coral_type) %>%
  summarize(maxn = max(n_total)) %>%
  mutate(maxn_cat = cut(maxn, breaks = c(0, 10, 100, 1000, 10000)))

dfs2 <- df %>%
  group_by(lat, lon, subregion, species, coral_type) %>%
  summarize(maxn = max(n_total)) %>%
  mutate(maxn_cat = cut(maxn, breaks = c(0, 10, 100, 1000, 10000)))

counts1 <- dfs %>%
  group_by(subregion) %>%
  summarize(maxtot1 = sum(maxn))

dfs %>%
  group_by(species, subregion) %>%
  summarize(maxtot1 = sum(maxn)) %>%
  arrange(subregion, species)

# Get sum treating each original data_provider.sitename as unique site, and get max observed at that site (on any date),
# and total by subregion
## This could overestimate, because diff data_providers could have surveyed same site/same corals, but
## here they'd be treated as different sites if sites named differently
counts2 <- df1 %>%
  group_by(site, subregion, species, coral_type) %>%
  summarize(maxn = max(n_total)) %>%
  group_by(subregion) %>%
  summarize(maxtot2 = sum(maxn))

counts <- full_join(counts1, counts2)

# Plot map of Florida
# Download satellite map for Florida
world <- rnaturalearth::ne_countries(scale = "large", returnclass = "sf")
# Create base map of Florida
basemap <- ggplot() +
  geom_sf(data = world, lwd = 0.1, fill = "gray90") +
  scale_fill_gradient2(high = "firebrick1", mid = "yellow", low = "forestgreen", 
                       midpoint = 0.5, limits = c(0, 1), na.value = NA,
                       labels = scales::label_percent(), name = "Corals\nbleached") +
  theme(text = element_text(size = 10),
        axis.title = element_blank(),
        panel.background = element_rect(fill = "lightsteelblue1"),
        panel.border = element_rect(colour = "black", fill=NA),
        panel.grid = element_blank(),
        legend.position = c(0.2, 0.5),
        legend.background = element_blank())


# Plot max number of ACER and APAL surveyed at each site
mapfig <- basemap +
  coord_sf(xlim = c(-83.2, -79.8), ylim = c(24.3, 26.5), expand = FALSE) +
  geom_point(data = dfs, aes(x = lon, y = lat, color = subregion, size = maxn_cat, shape = species), alpha = 0.4) +
  theme(legend.position = "none")

mapfig

# # # Surveys over time
ggplot(df, aes(x = date, fill = subregion)) +
  geom_histogram() +
  scale_x_date(date_breaks = "months", date_labels = "%b") +
  theme_classic() +
  labs(y = "Number of sites surveyed")

range(df$date)

# # Surveys by max DHW experienced anytime prior to survey
ggplot(df, aes(x = cummaxdhw, fill = subregion)) +
  geom_histogram(binwidth = 1, closed = "left") +
  theme_classic() +
  labs(y = "Number of sites surveyed")
```


# Spatial and temporal mortality patterns

```{r plot_mort, fig.width = 8, fig.height = 9}
# Plot Percent mortality over time
(mort.pts <- ggplot(df, aes(x = date, y = pct_mort, fill = subregion)) +
  geom_point(aes(size = sqrt(n_total), shape = coral_type), alpha = 0.4, stroke = 0.2) +
  scale_x_date(date_breaks = "months", date_labels = "%b") +
  scale_shape_manual(values = c(21, 22, 23)) +
  #scale_size_discrete(range = c(1,4)) +
  facet_grid(subregion~species) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Date", y = "Percent mortality"))
```

### Model mortality over time by species and subregion
```{r mod_mort, fig.width = 8, fig.height = 9}
# Fit model using glmer (sites as random factors)
# Rescale the date predictor
df$date_scaled <- scale(as.numeric(df$date))
mod.glmer <- glmer(cbind(n_dead, n_alive) ~ subregion * date_scaled * species + (1| siteclust/species),
                   family = binomial,
                   data = df)

# Get fitted values for glmer
res.glmer <- emmeans(mod.glmer, specs = c("date_scaled", "subregion", "species"), type = "response",
               at = list(date_scaled = seq(-2.168304, 2.380493, 0.014)))
datesd <- sd(as.numeric(df$date))
datemean <- mean(as.numeric(df$date))
res.glmer <- as.tibble(res.glmer) %>% mutate(date = as_date(date_scaled * datesd + datemean))

# Filter fitted values to just the range of observed data for each species/subregion
date_ranges <- df %>%
  group_by(species, subregion) %>%
  summarize(start = min(date), end = max(date))

res.glmer.f <- as.tibble(res.glmer) %>%
  group_by(species, subregion) %>%
  nest() %>%
  left_join(date_ranges) %>%
  mutate(data.f = pmap(list(data, start, end), function(data, start, end) data %>% filter(date >= start, date <= max(end, as.Date("2024-03-02"))))) %>%
  dplyr::select(subregion, species, data.f) %>%
  unnest(data.f)
```

```{r plot_mod_fits, fig.width = 8, fig.height = 9}
# Fitted Percent mortality on October 1, and April 1st from GLMER
glmer_fitted_mort <- res.glmer.f %>%
  ungroup() %>%
  mutate(date=ymd(date))%>%
  filter(date %in% c(ymd("2023-10-01"), ymd("2024-03-01"))) %>%
  dplyr::select(subregion, species, date, prob, asymp.LCL, asymp.UCL)

# Plot model fits over raw data, with panels arranged in vertical rows
(mort.fits.v <- mort.pts +
  geom_line(data = res.glmer.f, aes(y = prob), lty = 2) +
  geom_vline(xintercept = c(as.Date("2023-10-01"), as.Date("2024-03-01")), lwd = 0.25, color = "gray70") +
  geom_text(data = glmer_fitted_mort, aes(y = prob, label = scales::label_percent()(round(prob, 2))), fontface = "italic"))

# Another version of the figure with panels arranges in horizontal rows
# (mort.fits.h <- mort.fits.v + 
#   facet_grid(species~factor(subregion,
#       levels = rev(levels(subregion)))) +
#   theme(legend.position = "none") +
#   scale_y_continuous(labels = scales::percent) + 
#   labs(x = "", y = "Mortality"))

# Remove legend for saving
mort.fits.v2 <- mort.fits.v + theme(legend.position = "none")
ggsave(mort.fits.v2, filename = "outputs/FigS1.png", width = 150, height = 200, units = "mm")

# Make wide mortality table by subregion for each species
mortSpecTableWide <- glmer_fitted_mort %>%
  pivot_wider(id_cols = subregion, names_from = c(species, date), values_from = prob)

knitr::kable(mortSpecTableWide)
```

### Test for statistical differences between APAL and ACER
```{r apal_vs_acer}
# Overall difference between species across whole time series
# reduced model without species -- compare to full model with LRT for significance of species
mod.glmer.nospp <- glmer(cbind(n_dead, n_alive) ~ subregion * date_scaled + (1| siteclust:species),
                         family = binomial,
                         data = df)

anova(mod.glmer, mod.glmer.nospp, test = "Chisq")
BIC(mod.glmer, mod.glmer.nospp)
### There is an overall effect of species.
### Next, test if there is a difference in mortality between species at either of the two dates of interest

# Get date_scaled for October 1st 2023 and March 1st 2024
oct1_scaled <- df %>% filter(date == "2023-10-01") %>% pull(date_scaled) %>% unique(.)
mar1_scaled <- df %>% filter(date == "2024-03-01") %>% pull(date_scaled) %>% unique(.)

# Run pairwise test comparing species in each subregion on each date
emm <- emmeans(mod.glmer, specs = "species", by = c("subregion", "date_scaled"),
        at = list(date_scaled = c(oct1_scaled, mar1_scaled)))
rbind(contrast(emm, "pairwise"))
### No significant differences between species in any subregion on either date
### Note nearly significant difference in Biscayne on March 1st 2024; however, estimates for ACER on that date
### are poorly constrained because the latest observations were Nov 10th 2023... So, not concerned about this being a false negative.
```

### Estimate mortality across both species
```{r mod_allacro}
# Get results from glmer that includes species, with estimates averaged across species
# out.glmer <- emmeans(mod.glmer, specs = c("date_scaled", "subregion"), type = "response",
#                at = list(date_scaled = c(oct1_scaled, mar1_scaled))) %>% as_tibble() %>%
#              mutate(date = case_when(date_scaled == c(oct1_scaled) ~ as.Date("2023-10-01"),
#                                      date_scaled == c(mar1_scaled) ~ as.Date("2024-03-01")))

# Get results from glmer that does NOT include species as fixed factor
out.glmer.nospp <- emmeans(mod.glmer.nospp2, specs = c("date_scaled", "subregion"), type = "response",
               at = list(date_scaled = c(oct1_scaled, mar1_scaled))) %>% as_tibble() %>%
             mutate(date = case_when(date_scaled == c(oct1_scaled) ~ as.Date("2023-10-01"),
                                     date_scaled == c(mar1_scaled) ~ as.Date("2024-03-01")))


#### NEED TO CHOOSE WHICH RESULT TO USE
## result from glmer with species has massive confidence intervals (0-1) for MK and DRTO. seems problematic...
    ## this is because the confidence intervals for ACER (individual) are 0-1, be
## result from glmer without species has estimates that are not in between the species-specific estimates. is that a problem? Not necessarily -- just reflects differences in number of observations of diff species over diff time ranges. Biggest discrepancy is Biscayne, where lots of Acer was observed in summer/fall 2023, but not spring 2024, while more APAL was observed in 2024. In this case, fitting one model for all Acro independent of species may actually produce better estimates than the species-specific models, since they are both more data-poor. Going to go with the glmer without species, for this reason, and because it also produces interpretable confidence intervals, while the species-averaged estimates from the other model does not (for MK and DRTO, bc of Acer, see above).


## Make table for overall Acropora mortality for each subregion
out.glmer.nospp %>%
  dplyr::select(subregion, date, prob, asymp.LCL, asymp.UCL) %>%
  #pivot_wider(values_from = prob, names_from = date_scaled) %>%
  knitr::kable(caption = "Mortality from GLMER")

# # estimates across all subregions?
#  emmeans(mod.glmer.nospp, specs = c("date_scaled"), type = "response",
#                at = list(date_scaled = c(oct1_scaled, mar1_scaled))) %>% as_tibble() %>%
#              mutate(date = case_when(date_scaled == c(oct1_scaled) ~ as.Date("2023-10-01"),
#                                      date_scaled == c(mar1_scaled) ~ as.Date("2024-03-01")))
```

### Plot mortality across both species
```{r}
# Alter map extent to create space for adding graphs on top of map
mapfig2 <- basemap +
  geom_point(data = dfs2, aes(x = lon, y = lat, color = subregion, size = sqrt(maxn), shape = species), alpha = 0.4) +
  coord_sf(xlim = c(-83.5, -79), ylim = c(23.6, 26.5), expand = FALSE) +
  theme(axis.text.x = element_text(hjust = c(0.5, 0.5,0.5,0.5, 1))) +
  theme(legend.position = "none")

# # Create points to overlay on satellite map for USF presentation?
# justpoints <- ggplot(dfs2, aes(x = lon, y = lat, color = subregion, size = maxn_cat, shape = species)) +
#   geom_point(alpha = 0.8) +
#   theme_void() +
#   theme(legend.position = "none")
# ggsave(justpoints, filename = "justpoints.png", width = 207, height = 100, units = "mm")


# Create DHW plot to draw as inset on map
dhwmean <- read_csv("data/processed/MeanDHW_subregions.csv")
out <- read_csv("data/processed/DHW_siteclusts.csv")
new_dates <- seq.Date(from = as.Date("2024-01-01"), to = as.Date("2024-03-01"), by = "day")
new_data <- expand_grid(date = new_dates, subregion = levels(dhwmean$subregion), dhw = 0)
dhwmean <- bind_rows(dhwmean, new_data)
library(ggh4x)
dhwfig <- ggplot(out, aes(x = date, y = dhw, color = subregion)) +
  geom_line(aes(group = siteclust), lwd = 0.2, alpha = 0.5) +
  geom_line(data = dhwmean, lwd = 2, alpha = 0.6) +
  geom_segment(x = as.Date("2023-10-01"), xend = as.Date("2023-10-01"), y = -1, yend = 22, lty = 2, lwd = 0.1, color = "black") +
  geom_segment(x = as.Date("2024-03-01"), xend = as.Date("2024-03-01"), y = -1, yend = 9, lty = 2, lwd = 0.1, color = "black") +
  annotate("text", x = c(as.Date("2023-10-01"), as.Date("2024-03-01")), y = c(2, 2), 
           label = c(expression(italic(t)[1]), expression(italic(t)[2])), hjust = 1.3) +
  scale_x_date(breaks = as.Date(c("2023-06-01", "2023-10-01", "2024-03-01")), date_labels = "%b-%y",
               date_minor_breaks = "1 month", guide = "axis_minor") +
  labs(x = "", y = "Degree Heating Weeks (°C-weeks)") +
  theme_classic() +
  theme(legend.position = "none",
        axis.line = element_line(size = 0.1),
        axis.ticks = element_line(size = 0.1),
        axis.title.y = element_text(size = 8),
        plot.background = element_blank(),
        panel.background = element_blank(),
        ggh4x.axis.ticks.length.minor = rel(1))


# Create barplots of mortality at t1 and t2 for each subregion
bars <- out.glmer.nospp %>%
  dplyr::select(subregion, date, dead = prob, asymp.LCL, asymp.UCL) %>%
  mutate(alive = 1 - dead, subregion2 = subregion) %>%
  pivot_longer(c(dead, alive), names_to = "cat", values_to = "val") %>%
  left_join(counts1) %>%
  nest(data = -subregion) %>%
  mutate(barplots = map(data, ~ggplot(., aes(x = factor(date))) +
             geom_col(aes(y = val, fill = cat), color = "black", lwd = 0.1) +
             geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), lwd = 0.1, width = 0.1) +
             geom_text(aes(x = 1.5, y = 1, label = paste(subregion2, "\n", "n =", maxtot1), color = subregion2), vjust = -0.25, size = 2) +
             scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25), labels = c("0%", "", "50%", "", "100%")) +
             scale_x_discrete(labels = c(expression(italic(t)[1]), expression(italic(t)[2]))) +
             scale_color_manual(values = c("#F8766D","#B79F00","#00BA38","#00BFC4","#619CFF","#F564E3"), drop = FALSE) +
             scale_fill_manual(values = c("#a6d96a", "#de2d26")) +
             labs(x = "", y = "") +
             coord_cartesian(clip = 'off') +
             theme_classic() +
             theme(legend.position = "none", 
                   axis.line = element_line(size = 0.1),
                   axis.ticks = element_line(size = 0.1),
                   plot.background = element_blank(), 
                   panel.background = element_blank(),
                   axis.text.x = element_text(angle = 0, hjust = c(0.5, 0.5)),
                   axis.text.y = element_text(size = 5))))

bars$barplots[[6]] <- bars$barplots[[6]] +
  labs(y = expression(italic("Acropora")~spp.~mortality)) +
  theme(axis.title.y = element_text(size = 8))

# Plot mortality barplots on top of the map
outmapbars <- ggdraw() +
  draw_plot(mapfig2) +
  draw_plot(bars$barplots[[1]], height = 0.25, width = 0.2, x = 0.775, y = 0.65) +
  draw_plot(bars$barplots[[2]], height = 0.25, width = 0.2, x = 0.775, y = 0.41) +
  draw_plot(bars$barplots[[3]], height = 0.25, width = 0.2, x = 0.675, y = 0.18) +
  draw_plot(bars$barplots[[4]], height = 0.25, width = 0.2, x = 0.5, y = 0.075) +
  draw_plot(bars$barplots[[5]], height = 0.25, width = 0.2, x = 0.3, y = 0.05) +
  draw_plot(bars$barplots[[6]], height = 0.25, width = 0.2, x = 0.1, y = 0.05)

# Display map figure with additional plots
(outmap3 <- outmapbars +
  draw_plot(dhwfig, height = 0.48, width = 0.4, x = 0.1, y = 0.445))

# Save
ggsave("outputs/Fig2.png", plot = outmap3, width = 183, height = 140, units = "mm")
```


# Mortality rates with increasing heat stress

```{r plot_mort_dhw}
# Plot all data to visualize
ggplot(df, aes(x = cummaxdhw, y = pct_mort)) +
  facet_grid(species~coral_type) +
  geom_point()

ss <- filter(df, species == "APAL", cummaxdhw > 15)

ggplot(ss, aes(x = coral_type, y = pct_mort)) +
  # geom_violin() +
  # geom_boxplot() +
  geom_jitter()

filter(ss, coral_type == "Wild", pct_mort < 50)
```


### Compare *A. cervicornis* and *A. palmata* mortality response curves
```{r acer_vs_apal_mort_dhw}
# Fit the initial model
m3 <- drm(
  pct_mort ~ cummaxdhw, data = df, curveid = species, weights = n_total, type = "binomial",
  logDose = NULL, fct = LL.4(
    names = c("hill", "min", "max", "ED50"),
    fixed = c(NA, 0, 1, NA)
  )
)

# Get ED50s and ED95s
ed_values <- ED(m3, c(0.5, 0.95), type = "absolute", interval = "delta", level = 0.95, display = FALSE)
eds.res <- rownames_to_column(data.frame(ed_values), "par") %>% as_tibble() %>%
  separate(par, sep = ":", into = c("par", "species", "level"))

# Define new data for prediction
nd <- expand.grid(species = unique(df$species), cummaxdhw = seq(0, 23, 0.1))

# Number of bootstrap samples
n_boot <- 10000

# Bootstrap function for both predictions and ED comparisons
bootstrap_function <- function(iteration) {
  # Resample the data with replacement
  boot_data <- df[sample(1:nrow(df), replace = TRUE), ]
  
  # Refit the model on the bootstrap sample
  boot_model <- tryCatch({
    drm(
      pct_mort ~ cummaxdhw, data = boot_data, curveid = species, weights = n_total, type = "binomial",
      logDose = NULL, fct = LL.4(
        names = c("hill", "min", "max", "ED50"),
        fixed = c(NA, 0, 1, NA)
      )
    )
  }, error = function(e) NULL)  # Handle errors gracefully
  
  # Skip iteration if the model failed to converge
  if (is.null(boot_model)) return(list(preds = rep(NA, nrow(nd)), ED50_Diff = NA, ED95_Diff = NA))
  
  # Generate predictions for new data
  preds <- predict(boot_model, newdata = nd)
  
  # Extract ED50 and ED95 for both species
  ed_values <- as.tibble(ED(boot_model, c(50, 95), interval = "none", display = FALSE))$Estimate
  ed50_diff <- diff(ed_values[c(1,3)])  # ED50 difference
  ed95_diff <- diff(ed_values[c(2,4)])  # ED95 difference
  
  list(preds = preds, ED50_Diff = ed50_diff, ED95_Diff = ed95_diff)
}

# Run bootstrap
set.seed(123)  # For reproducibility
bootstrap_results <- replicate(n_boot, bootstrap_function(), simplify = FALSE)

# Extract bootstrap predictions and ED differences
bootstrap_preds <- do.call(cbind, lapply(bootstrap_results, `[[`, "preds"))
bootstrap_ed50_diff <- sapply(bootstrap_results, `[[`, "ED50_Diff")
bootstrap_ed95_diff <- sapply(bootstrap_results, `[[`, "ED95_Diff")

# Calculate prediction intervals
lower_PI <- apply(bootstrap_preds, 1, quantile, probs = 0.025, na.rm = TRUE)
upper_PI <- apply(bootstrap_preds, 1, quantile, probs = 0.975, na.rm = TRUE)
predicted <- predict(m3, newdata = nd)  # Original model predictions

# Statistical tests and confidence intervals for ED50 and ED95
summary_ed_results <- tibble(
  Mean_ED50_Diff = mean(bootstrap_ed50_diff, na.rm = TRUE),
  CI_Lower_ED50 = quantile(bootstrap_ed50_diff, 0.025, na.rm = TRUE),
  CI_Upper_ED50 = quantile(bootstrap_ed50_diff, 0.975, na.rm = TRUE),
  P_Value_ED50 = 2 * min(
    mean(bootstrap_ed50_diff >= 0, na.rm = TRUE), 
    mean(bootstrap_ed50_diff <= 0, na.rm = TRUE)
  ),
  
  Mean_ED95_Diff = mean(bootstrap_ed95_diff, na.rm = TRUE),
  CI_Lower_ED95 = quantile(bootstrap_ed95_diff, 0.025, na.rm = TRUE),
  CI_Upper_ED95 = quantile(bootstrap_ed95_diff, 0.975, na.rm = TRUE),
  P_Value_ED95 = 2 * min(
    mean(bootstrap_ed95_diff >= 0, na.rm = TRUE), 
    mean(bootstrap_ed95_diff <= 0, na.rm = TRUE)
  )
)

# Combine prediction results into a data frame
drc.res <- as_tibble(bind_cols(nd, fit = predicted, lower_PI = lower_PI, upper_PI = upper_PI))

# View results
print(summary_ed_results)

# Plot prediction intervals
ggplot(drc.res, aes(x = cummaxdhw, y = fit, color = species, fill = species)) +
  geom_line(size = 1) +
  geom_hline(yintercept = c(0.5, 0.95), linetype = 2) +
  geom_ribbon(aes(ymin = lower_PI, ymax = upper_PI), alpha = 0.2, linetype = 0) +
  labs(x = "DHW (°C-weeks)", y = "Mortality") +
  theme_classic() +
  scale_color_brewer(palette = "Set2", name = "", labels = c("A. cervicornis", "A. palmata")) +
  scale_fill_brewer(palette = "Set2", name = "", labels = c("A. cervicornis", "A. palmata")) +
  scale_y_continuous(labels = scales::label_percent()) +
  geom_text(
    data = eds.res, 
    aes(label = round(as.numeric(Estimate), 1), x = Estimate, y = as.numeric(level)), 
    show.legend = FALSE, 
    size = 4,
    nudge_x = c(3, 2, -1.5, -2),
    nudge_y = c(0.05, -0.05, 0.05, 0.05)
  ) +
  theme(
    legend.position = c(0.8, 0.15),
    legend.text = element_text(face = "italic")
  )

ggsave(filename = "outputs/Fig3.png", width = 89, height = 89, units = "mm")

# Plot bootstrap distributions for ED differences
bootstrap_ed_long <- tibble(
  ED50_Diff = bootstrap_ed50_diff,
  ED95_Diff = bootstrap_ed95_diff
) %>%
  pivot_longer(cols = everything(), names_to = "Parameter", values_to = "Difference")

ggplot(bootstrap_ed_long, aes(x = Difference, fill = Parameter)) +
  geom_histogram(bins = 30, color = "black", alpha = 0.7, position = "identity") +
  facet_wrap(~ Parameter, scales = "free", labeller = labeller(Parameter = c(
    ED50_Diff = "ED50 Difference",
    ED95_Diff = "ED95 Difference"
  ))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Bootstrap Distributions of ED50 and ED95 Differences between Species",
    x = "Difference",
    y = "Frequency"
  ) +
  theme_minimal()
```

### Compare wild vs. outplanted *A. palmata* mortality response curves

There was not enough data for wild *A. cervicornis* to make the same comparison for that species.

```{r apal_wild_outplant}
# For APAL, model outplant vs. wild
apal <- filter(df, species == "APAL", coral_type != "Wild/Outplant")
# Fit 4-parameter log-logistic dose-response curves for mortality as function of DHW, by species, with min=0 and max=1
apal.m3 <- drm(pct_mort ~ cummaxdhw, data = apal, curveid = coral_type, weights = n_total, type = "binomial",
         logDose = NULL, fct = LL.4(
           names = c("hill", "min", "max", "ED50"),
           ## Max parameter is fixed to 1, min is set to 0
           fixed = c(NA, 0, 1, NA)))

# Get ED50s and ED95s
ed_values <- ED(apal.m3, c(0.5, 0.95), type = "absolute", interval = "delta", level = 0.95, display = FALSE)
eds.res <- rownames_to_column(data.frame(ed_values), "par") %>% as_tibble() %>%
  separate(par, sep = ":", into = c("par", "coral_type", "level"))

# Number of bootstrap samples
n_boot <- 10000

# Define new data for prediction (same doses across coral types)
nd <- expand.grid(coral_type = unique(apal$coral_type), cummaxdhw = seq(0, 23, 0.1))

# Bootstrap function for predictions and ED comparisons
bootstrap_function <- function(iteration) {
  # Resample the data with replacement
  boot_data <- apal[sample(1:nrow(apal), replace = TRUE), ]
  
  # Refit the model on the bootstrap sample
  boot_model <- tryCatch({
      drm(
       pct_mort ~ cummaxdhw, data = boot_data, curveid = coral_type, weights = n_total, type = "binomial",
        logDose = NULL, fct = LL.4(
          names = c("hill", "min", "max", "ED50"),
          fixed = c(NA, 0, 1, NA)
        )
      )
  }, error = function(e) NULL)  # Handle errors gracefully
  
  # Skip iteration if the model failed to converge
  if (is.null(boot_model)) return(list(preds = rep(NA, nrow(nd)), ED50_Diff = NA, ED95_Diff = NA))
  
  # Generate predictions for new data
  preds <- predict(boot_model, newdata = nd)
  
  # Extract ED50 and ED95 for both coral types
  ed_values <- as.tibble(ED(boot_model, c(50, 95), interval = "none", display = FALSE))$Estimate
  ed50_diff <- diff(ed_values[c(1,3)])  # ED50 difference
  ed95_diff <- diff(ed_values[c(2,4)])  # ED95 difference
  
  list(preds = preds, ED50_Diff = ed50_diff, ED95_Diff = ed95_diff)
}

# Run bootstrap
set.seed(123)  # For reproducibility
bootstrap_results <- replicate(n_boot, bootstrap_function(), simplify = FALSE)

# Extract bootstrap predictions and ED differences
bootstrap_preds <- do.call(cbind, lapply(bootstrap_results, `[[`, "preds"))
bootstrap_ed50_diff <- sapply(bootstrap_results, `[[`, "ED50_Diff")
bootstrap_ed95_diff <- sapply(bootstrap_results, `[[`, "ED95_Diff")

# Calculate prediction intervals
lower_PI <- apply(bootstrap_preds, 1, quantile, probs = 0.025, na.rm = TRUE)
upper_PI <- apply(bootstrap_preds, 1, quantile, probs = 0.975, na.rm = TRUE)
predicted <- predict(apal.m3, newdata = nd)  # Original model predictions

# Statistical tests and confidence intervals for ED50 and ED95
summary_ed_results <- tibble(
  Mean_ED50_Diff = mean(bootstrap_ed50_diff, na.rm = TRUE),
  CI_Lower_ED50 = quantile(bootstrap_ed50_diff, 0.025, na.rm = TRUE),
  CI_Upper_ED50 = quantile(bootstrap_ed50_diff, 0.975, na.rm = TRUE),
  P_Value_ED50 = 2 * min(
    mean(bootstrap_ed50_diff >= 0, na.rm = TRUE), 
    mean(bootstrap_ed50_diff <= 0, na.rm = TRUE)
  ),
  
  Mean_ED95_Diff = mean(bootstrap_ed95_diff, na.rm = TRUE),
  CI_Lower_ED95 = quantile(bootstrap_ed95_diff, 0.025, na.rm = TRUE),
  CI_Upper_ED95 = quantile(bootstrap_ed95_diff, 0.975, na.rm = TRUE),
  P_Value_ED95 = 2 * min(
    mean(bootstrap_ed95_diff >= 0, na.rm = TRUE), 
    mean(bootstrap_ed95_diff <= 0, na.rm = TRUE)
  )
)

# Combine prediction results into a data frame
drc.res <- as_tibble(bind_cols(nd, fit = predicted, lower_PI = lower_PI, upper_PI = upper_PI))

# View ED summary results
print(summary_ed_results)

# Plot prediction intervals
ggplot(drc.res, aes(x = cummaxdhw, y = fit, color = coral_type, fill = coral_type)) +
  geom_line(size = 1) +
  geom_hline(yintercept = c(0.5, 0.95), linetype = 2) +
  geom_ribbon(aes(ymin = lower_PI, ymax = upper_PI), alpha = 0.2, linetype = 0) +
  labs(x = "DHW (°C-weeks)", y = "Mortality") +
  theme_classic() +
  scale_color_brewer(palette = "Set2", name = "", labels = c("Outplant", "Wild")) +
  scale_fill_brewer(palette = "Set2", name = "", labels = c("Outplant", "Wild")) +
  scale_y_continuous(labels = scales::label_percent()) +
  geom_text(
    data = eds.res,
    aes(label = round(as.numeric(Estimate), 1), x = pmin(Estimate, 23), y = as.numeric(level)),
    show.legend = FALSE,
    size = 4,
    nudge_x = c(-1, 0, 0.5, -0.5),
    nudge_y = c(0.05, 0.05, -0.05, -0.05)
  ) +
  theme(
    legend.position = c(0.8, 0.15)
  )

ggsave(filename = "outputs/FigS2.png", width = 89, height = 89, units = "mm")

# Plot bootstrap distributions for ED differences
bootstrap_ed_long <- tibble(
  ED50_Diff = bootstrap_ed50_diff,
  ED95_Diff = bootstrap_ed95_diff
) %>%
  pivot_longer(cols = everything(), names_to = "Parameter", values_to = "Difference")

ggplot(bootstrap_ed_long, aes(x = Difference, fill = Parameter)) +
  geom_histogram(bins = 30, color = "black", alpha = 0.7, position = "identity") +
  facet_wrap(~ Parameter, scales = "free", labeller = labeller(Parameter = c(
    ED50_Diff = "ED50 Difference (Wild - Outplant)",
    ED95_Diff = "ED95 Difference (Wild - Outplant)",
    scales = "free_x"
  ))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Bootstrap Distributions of ED50 and ED95 Differences",
    x = "Difference",
    y = "Frequency"
  ) +
  theme_minimal() +
  xlim(-20, 100)
```





