---
title: "Data tidying"
output: html_document
date: "2024-06-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(sp)
library(parzer)
library(lme4)
library(emmeans)
library(geosphere)
library(readxl)
library(tidyverse)
```

# Import data

## CRF data

```{r}
crf0 <- read_xlsx("CRF_FL_Acropora_data_.xlsx")

crf <- crf0 %>%
  mutate(
    data_provider = "crf",
    site = Site,
    lat = Latitude,
    lon = Longitude,
    date = as_date(Date),
    species = Spp,
    coral_type = `Wild/Outplant`,
    n_alive = `N (surviving colonies)`,
    n_dead = `N (colonies)` - `N (surviving colonies)`,
    n_geno = `N (Genotypes, if known)`
  ) %>%
  dplyr::select(data_provider, site, lat, lon, date, species, coral_type,
         n_alive, n_dead, n_geno)
```

## Mote data

```{r}
mml0 <- read_xlsx("FL_Acropora data_MoteMarineLaboratory.xlsx")

mml <- mml0 %>%
  mutate(
    data_provider = "mml",
    site = Site,
    lat = Latitude,
    lon = Longitude,
    date = as_date(Date),
    species = `Spp (APAL/ACER)`,
    coral_type = `Wild/Outplant`,
    n_alive = `N (surviving colonies)`,
    n_dead = `N (colonies)` - `N (surviving colonies)`,
    n_geno = `N (Genotypes, if known)`
  ) %>%
  dplyr::select(data_provider, site, lat, lon, date, species, coral_type,
         n_alive, n_dead, n_geno)
```

## NSU data

```{r}
nsu0 <- read_xlsx("NSU_Gilliam_Acropora data.xlsx")

nsu <- nsu0 %>%
  mutate(
    data_provider = "nsu",
    site = Site,
    lat = parse_lat(Latitude),
    lon = parse_lon(Longitude) * -1,
    date = as_date(Date),
    species = `Spp (APAL/ACER)`,
    coral_type = `Wild/Outplant`,
    n_alive = `N (surviving colonies)`,
    n_dead = `N (colonies)` - `N (surviving colonies)`,
    n_geno = `N (Genotypes, if known)`
  ) %>%
  dplyr::select(data_provider, site, lat, lon, date, species, coral_type,
         n_alive, n_dead, n_geno)
```

## UM data

```{r}
um0 <- read_xlsx("UM_AcroporaData.xlsx")

um <- um0 %>%
  mutate(
    data_provider = "um",
    site = Site,
    lat = Latitude,
    lon = Longitude,
    date = as_date(Date),
    species = `Spp (APAL/ACER)`,
    coral_type = `Wild/Outplant`,
    n_alive = `N (surviving colonies)`,
    n_dead = `N (colonies)` - `N (surviving colonies)`,
    n_geno = `N (Genotypes, if known)`
  ) %>%
  dplyr::select(data_provider, site, lat, lon, date, species, coral_type,
         n_alive, n_dead, n_geno)
```

## FWC data

```{r}
fwc0 <- read_xlsx("FL_APAL_Summary Data.xlsx", sheet = "FWC", 
                  n_max = 28,
                  na = c("N/A", "n/a")) %>%
  filter(Site != "REGION SUMMARY")

fwc <- fwc0 %>%
  mutate(
    data_provider = "fwc",
    site = Site,
    lat = Latitude,
    lon = Longitude,
    date = as_date(Date),
    species = `Spp`,
    coral_type = `Wild/Outplant`,
    n_alive = `N (surviving colonies)`,
    n_dead = `N (colonies)` - `N (surviving colonies)`,
    n_geno = `N (Genotypes, if known)`
  ) %>%
  dplyr::select(data_provider, site, lat, lon, date, species, coral_type,
         n_alive, n_dead, n_geno)
```

## USGS data

```{r}
usgs0 <- read_xlsx("FL_APAL_Summary Data.xlsx", sheet = "KUFFNER", 
                  n_max = 12,
                  na = c("N/A", "n/a")) %>%
  filter(Site != "REGION SUMMARY")

usgs <- usgs0 %>%
  mutate(
    data_provider = "usgs",
    site = Site,
    lat = Latitude,
    lon = Longitude,
    date = as_date(Date),
    species = `Spp`,
    coral_type = `Wild/Outplant`,
    n_alive = `N (surviving colonies)`,
    n_dead = `N (colonies)` - `N (surviving colonies)`,
    n_geno = `N (Genotypes, if known)`
  ) %>%
  dplyr::select(data_provider, site, lat, lon, date, species, coral_type,
         n_alive, n_dead, n_geno)
```

## DW data

```{r}
dw0 <- read_xlsx("FL_APAL_Summary Data_DEW2.xlsx", sheet = "DANA W", 
                  n_max = 95,
                  na = c("N/A", "n/a")) %>%
  filter(Site != "REGION SUMMARY")

dw <- dw0 %>%
  mutate(
    data_provider = "dw",
    site = Site,
    lat = Latitude,
    lon = Longitude,
    date = as_date(Date),
    species = `Spp`,
    coral_type = `Wild/Outplant`,
    n_alive = `N (surviving colonies)`,
    n_dead = `N (colonies)` - `N (surviving colonies)`,
    n_geno = `N (Genotypes, if known)`
  ) %>%
  dplyr::select(data_provider, site, lat, lon, date, species, coral_type,
         n_alive, n_dead, n_geno)

# Dana counted fission/partial mortality resulting in separate ramets, so number of survivors is sometimes higher than original total number of colonies. In these cases, set number of survivors to equal the number of original colonies.
dw <- dw %>%
  mutate(n_dead = pmax(0, n_dead))

dw <- dw %>%
  group_by(site, date, species, coral_type) %>%
  summarize(n_alive = sum(n_alive),
            n_dead = sum(n_dead),
            lat = mean(lat),
            lon = mean(lon))
```

# Shedd DRTO ACER data
```{r}
# June outplants
jun <- expand_grid(date = as.Date("2023-06-30"),
                   site = c("Pulaski Shoal", "Texas Rock", "Southwest Channel"),
                   n_dead = 0,
                   n_alive = 180,
                   coral_type = "Outplant",
                   data_provider = "shedd",
                   species = "ACER")

dtac <- read_csv("fouling_metadata.csv")

dtac <- dtac %>%
  filter(!is.na(Ranking)) %>%
  group_by(Site, Lat, Lon) %>%
  summarize(n_dead = sum(Ranking > 1),
            n_alive = sum(Ranking == 1)) %>%
  mutate(date = as.Date("2023-09-09"),
         data_provider = "shedd",
         species = "ACER",
         coral_type = "Outplant",
         n_geno = NA) %>%
  ungroup() %>%
  dplyr::select(data_provider, site = Site, lat = Lat, lon = Lon,
                date, species, coral_type, n_alive, n_dead, n_geno)
  
dtac <- dtac %>%
  bind_rows(jun) %>%
  group_by(site) %>%
  mutate(lat = ifelse(is.na(lat), first(na.omit(lat)), lat),
         lon = ifelse(is.na(lon), first(na.omit(lon)), lon)) %>%
  ungroup()
```

# MIR data
```{r}
mir0 <- read_xlsx("MIR_Acropora data.xlsx")

mir <- mir0 %>%
  mutate(
    data_provider = "mir",
    site = Site,
    lat = Latitude,
    lon = Longitude,
    date = as_date(Date),
    species = `Spp (APAL/ACER)`,
    coral_type = `Wild/Outplant`,
    n_alive = `N (surviving colonies)`,
    n_dead = `N (colonies)` - `N (surviving colonies)`,
    n_geno = `N (Genotypes, if known)`
  ) %>%
  dplyr::select(data_provider, site, lat, lon, date, species, coral_type,
         n_alive, n_dead, n_geno)
```

# Combine all data, 
```{r}
df0 <- bind_rows(
  crf, dw, fwc, mml, nsu, um, usgs, dtac, mir
) %>%
  mutate(n_total = n_dead + n_alive,
         pct_mort = n_dead / n_total)
```

# cluster sites, assign subregions
```{r}
# Ensure unique site names (some diff. groups used same site name (e.g., "100") for diff. sites)
# Generate distinct site names
df <- df0 %>%
  group_by(site) %>%
  mutate(site_count = n_distinct(paste(lat, lon))) %>%
  group_by(site, lat, lon) %>%
  mutate(site = ifelse(site_count == 1, site, paste0(site, "_", cur_group_id()))) %>%
  ungroup()

# Cluster sites that are close together
dist_mat <- distm(df[, c("lon", "lat")], fun = distHaversine)
hc <- hclust(as.dist(dist_mat), method = "complete")
clusters <- cutree(hc, h = 2000)  # Cluster based on 2000m threshold
df$siteclust <- as.character(clusters)

# ggplot(df, aes(x = lon, y = lat, color = as.factor(siteclust))) +
#   geom_point()


# Combine all data by site cluster/date
df <- df %>%
  group_by(siteclust) %>%
  mutate(lat = mean(lat),
         lon = mean(lon)) %>%
  group_by(siteclust, date, species, coral_type, lat, lon) %>%
  summarize(n_alive = sum(n_alive),
            n_dead = sum(n_dead),
            n_total = n_dead + n_alive,
            pct_mort = n_dead / n_total) %>%
  ungroup()

# Assign Subregions based on DRM subregions
load("~/Projects/DRM2024/output/2023.RData")
subregions <- distinct(df23, Site, Latitude, Longitude, Subregion) %>%
  dplyr::select(site = Site, lat = Latitude, lon = Longitude, subregion = Subregion) %>%
  drop_na(lat)

# Calculate distances between acropora data locations and drm sites (with subregion info)

# Omit acropora data sites with missing coords
df <- df %>% drop_na(lat, lon)
dist_matrix <- distm(df[, c("lat", "lon")], subregions[, c("lat", "lon")], fun = distHaversine)

# Find the nearest location in drm data (with subregion) for each entry in acropora data
nearest_indices <- apply(dist_matrix, 1, which.min)

# Add nearest subregion to df
df <- df %>%
  mutate(subregion = subregions$subregion[nearest_indices]) %>%
  drop_na(date) %>%
  droplevels()
```




# Import DHW data
```{r}
# Import all NetCDF files for each date from June 1 through December 31, 2023
nc.files <- list.files("~/Projects/DRM2024/data/dhw/2023", pattern = "*.nc$", full.names = TRUE)

# Get coordinate box for Florida Reef Tract
#nc <- nc_open("data/dhw/ct5km_dhw_v3.1_20230901.nc")
#nct <- ncvar_get(nc, "degree_heating_week", start = c(2000, 1240,1), count = c(21, 81,1))
# Longitude -83.525 --> index 1930
#nc$dim$lon$vals[2021]
# Longitude -78.975 --> index 2021
# Latitude 24 --> index 1321
#nc$dim$lat$vals[1240]
# Latitude 28 --> index 1240

# Get DRM survey site coordinates
# Import data
sites <- df %>%
  distinct(siteclust, subregion, lon, lat) %>%
  mutate(subregion = factor(subregion, levels = rev(levels(subregion))))

# Open all DHW NCDF4 files
library(ncdf4) # package for netcdf manipulation
library(raster) # package for raster manipulation
nc.data <- tibble(file = nc.files) %>%
  mutate(date = map_chr(file, ~str_extract(., "2023\\d+"), format = "%Y%m%d"),
         data = map(file, nc_open),
      dhw.all = map(data, ~ncvar_get(., "degree_heating_week", start = c(1930, 1240, 1), count = c(91, 81, 1))),
            r = map(dhw.all, ~raster(t(.), xmn = -83.525, xmx = -78.975, ymn = 23.975, ymx = 28.025,
                                     crs = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs+ towgs84=0,0,0"))),
       coords = list(sites),
          dhw = map2(r, coords, ~raster::extract(.x, SpatialPoints(.y[,3:4]), method = "simple")))

out <- nc.data %>% dplyr::select(date, coords, dhw) %>% unnest() %>%
  mutate(date = as_date(date, format = "%Y%m%d"))

# Check that all sites got DHW values
dhws <- out %>%
  group_by(siteclust) %>%
  dplyr::summarize(avgdhw = mean(dhw))
nodhw <- dhws %>% filter(is.na(avgdhw))
#nodhw
## 8 siteclusts did not get DHW -- outside of 5km pixels (too close to land)

# Assign missing dhws value of closest site that has a dhw value
library(FastKNN)

# matrix of neighbours
n <- nrow(sites)
k <- 35 # Number of nearest neighbors to find
dm <- dist(sites[,3:4])
nn <- matrix(0,n,k) # n x k nearest neighbors matrix
for (i in 1:n) {
   nn[i,] = k.nearest.neighbors(i, as.matrix(dm), k = k)
}
nn[] <- as.character(sites$siteclust)[c(nn)]

# Get names of all sites and their nearest neighbors
neighbors <- as_tibble(cbind(siteclust = as.character(sites$siteclust), nn))
# Get dhws of all sites and dhws of their nearest neighbors
neighborsdhw <- neighbors %>%
  pivot_longer(-1, names_to = "nno", values_to = "nname") %>%
  mutate(nno = parse_number(nno) - 1) %>%
  mutate(navgdhw = dhws$avgdhw[match(nname, dhws$siteclust)])

finddhw <- neighborsdhw %>%
  filter(siteclust %in% nodhw$siteclust)

dhwmatch <- finddhw %>% 
  group_by(siteclust) %>%
  filter(!is.na(navgdhw)) %>%
  dplyr::summarize(closest_n_with_dhw = nname[nno == min(nno)]) %>%
  right_join(nodhw) %>%
  dplyr::select(1, 2)# %>%
  #print(n = nrow(.))
  
# Output matches
matches <- left_join(dhwmatch, sites)
#range(matches$Longitude)
#range(matches$Latitude)
matches2 <- left_join(dplyr::select(matches, siteclust = closest_n_with_dhw), sites) %>%
  distinct(siteclust, lon, lat)
# 
# # Plot matches
# library(ggrepel)
# ggplot(matches, aes(x = lon, y = lat)) +
#   geom_point() +
#   geom_point(data = matches2, color = "green") +
#   geom_label_repel(data = matches2, aes(label = site), color = "green", max.overlaps = 100, force = 80) +
#   geom_label_repel(aes(label = site), max.overlaps = 100, force = 80) +
#   xlim(-80.125, -80) +
#   ylim(26, 26.75)

# Replace missing dhws with values from matching sites
#matches
fixing <- out %>%
  nest(data = c(date, dhw)) 

fixed <- fixing
### dhw data for sites where missing gets replaced with dhw data for the site of its closest neighbor site with dhw data
fixed[match(matches$siteclust, fixing$siteclust), "data"] <- fixing[match(matches$closest_n_with_dhw, fixing$siteclust), "data"]

# ### test if it worked (dhw was missing for site 100, should have been replaced with dhw data from KC)
# fixing %>% filter(site == "100") %>% pull(data)
# fixed %>% filter(site == "KC") %>% pull(data)
# 
# identical(fixed$data[fixed$site == "100"], fixed$data[fixed$site == "KC"])
# y1 <- fixed$data[fixed$site == "100"][[1]]
# y2 <- fixed$data[fixed$site == "KC"][[1]]
# cbind(y1, y2)

# Unnest fixed output
outfixed <- fixed %>% unnest()
#outfixed %>% filter(is.na(dhw))
dhws <- outfixed %>% dplyr::select(siteclust, lon, lat, date, dhw)

# We only downloaded DHW from June through December 2024, but we have Acro survey data from before and after that interval. Those surveys would have DHW values of zero, so add those in...
df1 <- expand_grid(
  date = seq(as.Date("2023-03-14"), as.Date("2024-05-10"), by = "days"),
  dhws %>% distinct(siteclust, lat, lon)
)
dhws <- full_join(dhws, df1) %>%
  arrange(siteclust, lon, lat, date) %>%
  replace_na(list(dhw = 0))


# Get max. DHW value for on OR BEFORE date surveyed
dhws <- dhws %>%
  group_by(siteclust) %>%
  mutate(cummaxdhw = cummax(dhw))


# For each survey, join DHW on date of survey, and max DHW experienced on or BEFORE survey
df <- left_join(df, dhws, by = c("siteclust", "lon", "lat", "date")) %>%
  mutate(subregion = factor(subregion, levels = levels(sites$subregion)))
```

# add n_total_cat
```{r}
df <- df %>%
  mutate(n_total_cat = cut(n_total, breaks = c(0, 10, 100, 1000, 7000), right = FALSE))

df %>%
  count(n_total_cat)
```

# Surveys over space and time

```{r}
# # Map sites
# ggplot(df, aes(x = lon, y = lat, color = subregion)) +
#   geom_point(aes(size = n_total), alpha = 0.3) +
#   scale_size(range = c(1, 10)) +
#   theme_classic()
# 
# # Surveys over time
# ggplot(df, aes(x = date, fill = subregion)) +
#   geom_histogram() +
#   scale_x_date(date_breaks = "months", date_labels = "%b") +
#   theme_classic() +
#   labs(y = "Number of sites surveyed")

# # Surveys by max DHW experienced anytime prior to survey
# ggplot(df, aes(x = cummaxdhw, fill = subregion)) +
#   geom_histogram() +
#   theme_classic() +
#   labs(y = "Number of sites surveyed")

# Filter to just June onward...
df <- df %>% filter(date > "2023-06-01")
```

# Mortality over time by subregion

```{r, fig.width = 8, fig.height = 9}
# Plot Percent mortality over time
(mort <- ggplot(df, aes(x = date, y = pct_mort, color = subregion)) +
  geom_point(aes(size = n_total_cat, shape = coral_type), alpha = 0.4) +
  scale_x_date(date_breaks = "months", date_labels = "%b") +
  scale_size_discrete(range = c(1,4)) +
  facet_grid(subregion~species) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Date", y = "Percent mortality"))

# Combine Mid-Upper Keys Transition with Middle Keys (only ~2 timepoints)
df <- df %>%
  mutate(subregion = case_when(subregion == "Mid-Upper Keys Transition" ~ "Middle Keys",
                               TRUE ~ subregion)) %>%
  mutate(subregion = factor(subregion, levels = c("Broward-Miami", "Biscayne", "Upper Keys",
                                                  "Middle Keys", "Lower Keys", "Tortugas--Dry Tortugas NP")))

(mort <- mort %+% df)
```

# Fit models

```{r, fig.width = 8, fig.height = 9}
# Fit model using glm
mod.glm <- glm(cbind(n_dead, n_alive) ~ subregion * date * species, 
               family = "binomial", data = df)
mod.glm2 <- glm(cbind(n_dead, n_alive) ~ subregion + date + species + subregion:date:species, 
                family = "binomial", data = df)
AIC(mod.glm, mod.glm2)

# Fit model using glmer (sites as random factors)
# Rescale the date predictor
df$date_scaled <- scale(as.numeric(df$date))
mod.glmer <- glmer(cbind(n_dead, n_alive) ~ subregion * date_scaled * species + (1| siteclust/species),
                   family = binomial,
                   data = df)
mod.glmer2 <- glmer(cbind(n_dead, n_alive) ~ subregion + date_scaled + species + 
                    subregion:date_scaled:species + (1| siteclust/species),
                    family = binomial,
                    data = df)
AIC(mod.glmer, mod.glmer2)

# Get fitted values for glm
res.glm <- emmeans(mod.glm, specs = c("date", "subregion", "species"), type = "response",
               at = list(date = seq(min(df$date), max(df$date), by = "days")))
res.glm <- as.tibble(res.glm)

# Get fitted values for glmer
res.glmer <- emmeans(mod.glmer, specs = c("date_scaled", "subregion", "species"), type = "response",
               at = list(date_scaled = seq(-2.168304, 2.380493, 0.015)))
datesd <- sd(as.numeric(df$date))
datemean <- mean(as.numeric(df$date))
res.glmer <- as.tibble(res.glmer) %>% mutate(date = as.Date(date_scaled * datesd + datemean))

# Filter fitted values to just the range of observed data for each species/subregion
date_ranges <- df %>%
  group_by(species, subregion) %>%
  summarize(start = min(date), end = max(date))

res.glm.f <- as.tibble(res.glm) %>%
  group_by(species, subregion) %>%
  nest() %>%
  left_join(date_ranges) %>%
  mutate(data.f = pmap(list(data, start, end), function(data, start, end) data %>% filter(date > start, date < end))) %>%
  dplyr::select(subregion, species, data.f) %>%
  unnest(data.f)

res.glmer.f <- as.tibble(res.glmer) %>%
  group_by(species, subregion) %>%
  nest() %>%
  left_join(date_ranges) %>%
  mutate(data.f = pmap(list(data, start, end), function(data, start, end) data %>% filter(date > start, date < end))) %>%
  dplyr::select(subregion, species, data.f) %>%
  unnest(data.f)
```

# Plot
```{r, fig.width = 8, fig.height = 9}
# Fitted Percent mortality on October 1, and April 1st from GLM
glm_fitted_mort <- res.glm.f %>%
  ungroup() %>%
  filter(date %in% c("2023-10-01", "2024-01-01", "2024-04-01")) %>%
  dplyr::select(subregion, species, date, prob)

glmer_fitted_mort <- res.glmer.f %>%
  ungroup() %>%
  filter(date %in% c("2023-10-01", "2024-01-01", "2024-04-01")) %>%
  dplyr::select(subregion, species, date, prob)

# Plot fitted lines over raw data
mort +
  geom_line(data = res.glm.f, aes(y = prob)) +
  geom_line(data = res.glmer.f, aes(y = prob), lty = 2) +
  geom_vline(xintercept = c(as.Date("2023-10-01"), as.Date("2024-01-01"), as.Date("2024-04-01")), lwd = 0.25, color = "gray70") +
  geom_text(data = glm_fitted_mort, aes(y = prob, label = scales::label_percent()(round(prob, 2)))) +
  geom_text(data = glmer_fitted_mort, aes(y = prob, label = scales::label_percent()(round(prob, 2))), fontface = "italic")

# ggplot(as.tibble(res), aes(x = date, y = prob)) +
#   geom_line() +
#   facet_grid(species ~ subregion)
```


# Any difference between cervicornis and palmata?
```{r}
# On October 1st, 2023
emm1001 <- emmeans(mod.glm, specs = "species", by = c("date", "subregion"), type = "response",
               at = list(date = as.Date("2023-10-01")))
rbind(contrast(emm1001, "pairwise"))

# On April 1st, 2024
emm0401 <- emmeans(mod, specs = "species", by = c("date", "subregion"), type = "response",
               at = list(date = as.Date("2024-04-01")))
rbind(contrast(emm0401, "pairwise"))

# species is significant.... effect sizes aren't that big though.

# Overall effect of species
# reduced model without species -- compare to full model with LRT for significance of species
mod.glm.nospp <- glm(formula = cbind(n_dead, n_alive) ~ subregion * date, 
                     family = "binomial", data = df)
anova(mod.glm, mod.glm.nospp, test = "Chisq")
BIC(mod.glm, mod.glm.nospp)

mod.glmer.nospp <- glmer(cbind(n_dead, n_alive) ~ subregion * date_scaled + (1| siteclust/species),
                         family = binomial,
                         data = df)
anova(mod.glmer, mod.glmer.nospp, test = "Chisq")
BIC(mod.glmer, mod.glmer.nospp)
```

# even though species is significant, model all acropora together
```{r}
# Fit model using glm
mod.all.glm <- glm(cbind(n_dead, n_alive) ~ subregion * date, 
               family = "binomial", data = df)

# Fit model using glmer (sites as random factors)
mod.all.glmer <- glmer(cbind(n_dead, n_alive) ~ subregion * date_scaled + (1|siteclust/species),
                       family = binomial, data = df)

# Fitted values for glm
res.all.glm <- emmeans(mod.all.glm, specs = c("date", "subregion"), type = "response",
                       at = list(date = seq(as.Date("2023-03-14"), as.Date("2024-06-01"), by = "day")))

# Fitted values for glmer
res.all.glmer <- emmeans(mod.all.glmer, specs = c("date_scaled", "subregion"), type = "response",
                         at = list(date_scaled = seq(-2.699, 2.167, 0.015)))

# Filter fitted values to just the range of observed data
date_ranges <- df %>%
  group_by(subregion) %>%
  summarize(start = min(date), end = max(date))
res.all.glm.f <- as.tibble(res.all.glm) %>%
  group_by(subregion) %>%
  nest() %>%
  left_join(date_ranges) %>%
  mutate(data.f = pmap(list(data, start, end), function(data, start, end) data %>% filter(date > start))) %>% #, date < end))) %>%
  dplyr::select(subregion, data.f) %>%
  unnest(data.f)

date_ranges_scaled <- df %>%
  group_by(subregion) %>%
  summarize(start = min(date_scaled), end = max(date_scaled))
res.all.glmer.f <- as.tibble(res.all.glmer) %>%
  group_by(subregion) %>%
  nest() %>%
  left_join(date_ranges_scaled) %>%
  mutate(data.f = pmap(list(data, start, end), function(data, start, end) data %>% filter(date_scaled > start))) %>% #, date < end))) %>%
  dplyr::select(subregion, data.f) %>%
  unnest(data.f) %>%
  mutate(date = as.Date(date_scaled * datesd + datemean))

as.tibble(res.all.glm.f) %>%
  ggplot(aes(x = date, y = prob, color = subregion)) +
  geom_line() +
  geom_line(data = as.tibble(res.all.glmer.f), lty = 2) +
  scale_x_date(date_breaks = "months", date_labels = "%b") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = "Date", y = "Percent mortality")

# Pull out Oct1 and Apr1 mortality for each region using GLM
as.tibble(res.all.glm.f) %>%
  filter(date %in% c("2023-10-01", "2024-04-01")) %>%
  ggplot(aes(x = factor(date), y = prob)) +
  geom_col() +
  facet_wrap(~subregion, ncol = 1)

# GLMER
as.tibble(res.all.glmer.f) %>%
  filter(date %in% c("2023-10-01", "2024-04-01")) %>%
  ggplot(aes(x = factor(date), y = prob)) +
  geom_col() +
  facet_wrap(~subregion, ncol = 1)
```


```{r}

mod <- glm(cbind(n_dead, n_alive) ~ cummaxdhw * species, family = "binomial", data = df)
res <- emmeans(mod, specs = "cummaxdhw", by = "species", type = "response", at = list(cummaxdhw = seq(0, 22, 0.1)))


# Percent mortality w/ maxDHW
ggplot(df, aes(x = cummaxdhw, y = pct_mort)) +
  geom_point(aes(size = n_total_cat, shape = species, color = subregion), alpha = 0.4) +
  scale_size_discrete(range = c(1,4)) +
  theme_classic() +
  geom_line(data = as_tibble(res), aes(y = prob, lty = species))
```

